---
title: 'Week 8 In-Class'
author: "Mina Mehdinia"
date: '2022-05-29'
output: html_document
---

```{r}

library(readr)
library(tidyr)
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(pander)
library(nycflights13)
library(gapminder)
```

##  <span style="color:cornflowerblue">In-class problem</span> {.smaller .build}

Filter from the `restaurants` data set the inspections for coffee shops only. Let's say that a coffee shop is anything that has "COFFEE", "ESPRESSO", or "ROASTER" in the name. 

```{r, echo=T, warning=FALSE, message=FALSE}
library(tidyverse)
load("data/restaurants_sub.Rdata")

coffee_data <- restaurants %>% 
  mutate(Name = str_to_upper(Name)) %>% # capitalize Name column
  filter(str_detect(Name, "COFFEE|ESPRESSO|ROASTER")) 

coffee_data %>% 
  distinct(Name) %>%  # remove duplicate
  head() # see first few rows




```

## <span style="color:cornflowerblue">In-class problems</span> {.smaller .build}

For the following vector of randomly generated names, write a regular expression that,
```{r}
thenames <- c("Jeremy Cruz", "Nathaniel Le", "Jasmine Chu", "Bradley Calderon Raygoza", 
"Quinten Weller", "Katelien Kanamu-Hauanio", "Zuhriyaa al-Amen", 
"Travale York", "Alexis Ahmed", "David Alcocer", "Jairo Martinez", 
"Dwone Gallegos", "Amanda Sherwood", "Hadiyya el-Eid", "Shaimaaa al-Can", 
"Sarah Love", "Shelby Villano", "Sundus al-Hashmi", "Dyani Loving", 
"Shanelle Douglas")
```

* detects if the person's first name starts with a vowel (a,e,i,o,u)

```{r}
thenames[str_detect(thenames, "^[aeiouAEIOU]")] 
```

* detects if the person's last name starts with a vowel

```{r}
thenames[str_detect(thenames, " [aeiouAEIOU]")]
```


* detects if either the person's first or last name start with a vowel
```{r}
thenames[str_detect(thenames, "^[aeiouAEIOU]|\\s\\b[aeiouAEIOU]")]
```


* detects if neither the person's first nor last name start with a vowel

```{r}
thenames[str_detect(thenames, "^[^AEIOU].* [^aeiouAEIOU]-?[^AEIOU]")] 

```


## <span style="color:cornflowerblue">In-class problem</span> {.smaller .build}

```{r}
text = c(
  "apple",
  "219 733 8965",
  "329-293-8753",
  "Work: (579) 499-7527; Home: (543) 355 3679"
)
```

* Write a regular expression that will extract *all* phone numbers contained in the vector above.
```{r}
patt1 <- "\\(?\\d{3}\\)?-? ?\\d{3}-? ?\\d{4}"

str_extract_all(text, patt1)
```


* Once that works use groups to extract the area code separately from the rest of the phone number.
```{r}
str_extract(unlist(str_extract_all(text, patt1)),"(\\d{3})") 
```


## <span style="color:cornflowerblue">In-class problems</span> {.smaller .build}

Use the appropriate lubridate function to parse each of the following dates:

```{r}
library(lubridate)
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014

mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)

```



## <span style="color:cornflowerblue">In-class problem</span> {.smaller .build}

1. On what day of the week should you leave if you want to minimise the chance of a delay?
```{r}
library(nycflights13)
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time")) 

flights_dt %>%  
  mutate(dow = wday(dep_time, label = T)) %>% 
  group_by(dow) %>% 
  filter(!is.na(arr_delay)) %>% 
  select(arr_delay, dow) %>% 
  summarise(avg_delay = median(arr_delay, na.rm = T), n = n()) %>% 
  ggplot(aes(dow, avg_delay)) +
  geom_bar(stat = "identity", fill = "red") +
  xlab("day of the week") + ylab("median delay (mins)") +
  ggthemes::theme_economist()
```


2. Confirm the hypothesis that the earlier arrivals of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.

```{r}
flights_dt %>% 
  filter(!is.na(arr_delay)) %>% 
  select(arr_delay, dep_time, sched_dep_time) %>% 
  mutate(early_dep = factor(ifelse(dep_time < sched_dep_time, 1, 0),
                            levels = c(1,0),
                            labels = c("early departure", "on-time/late departure")),
         minute = minute(dep_time)) %>% 
  group_by(minute, early_dep) %>% 
  summarise(avg_delay = mean(arr_delay, na.rm = T),
            n = n(),
            .groups = 'drop') %>% # remove warning msg 
  
  ggplot(aes(x = minute, y = n)) +
  geom_bar(stat = "identity", aes(fill = early_dep)) +
  facet_wrap(early_dep~.) +
  xlab("departure minute") +
  ylab("number of flights") +
  ggthemes::theme_economist_white(gray_bg = F) +
  theme(legend.position = "none")
```
