---
title: "In Class Problem"
author: "Mina Mehdinia"
date: '2022-05-15'
output: html_document
---
```{r}
library(nycflights13)
library(dplyr)
library(ggplot2) 
library(tidyverse)
```

# Tuesday 
Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

You might want to use the size or colour of the points to display the average delay for each airport.

```{r}
data(flights)
data(airports)
data(planes)

```

```{r}
flights %>% 
  group_by(dest) %>% 
  filter(!is.na(arr_delay)) %>% 
  summarise(mean_delay = mean(arr_delay)) %>% 
  left_join(airports, by = c("dest" = "faa")) %>% 
  filter(!is.na(lat) & !is.na(lon)) %>% #remove Na's
  filter(lon > (-140)) %>% 
  mutate(mean_delay = ifelse(mean_delay < 0, 0, mean_delay)) %>% 
  ggplot(aes(x = lon, y = lat, col = mean_delay, size = mean_delay)) +
  borders("state") + 
  geom_point() + 
  coord_quickmap() + 
  xlab("Longitude") + 
  ylab("Latitude")
```
Add the location of the origin and destination (i.e. the lat and lon) to flights.
```{r}
flights %>% 
  left_join(airports %>% select(faa, lat, lon),
            by = c("origin" = "faa")) %>%
  left_join(airports %>% select(faa,lat,lon),
            by = c("dest" = "faa"), suffix = c("", " .dest"))
```
Is there a relationship between the age of a plane and its delays?
```{r}
flights %>% select(year, month, day, carrier, tailnum, dep_delay, arr_delay) %>% 
  left_join(planes %>% select(tailnum, year),
            by = "tailnum", suffix = c("", "_built")) %>% 
  mutate(plane_age = (year - year_built)) %>% 
  filter(!is.na(plane_age) & !is.na(dep_delay) &
           !is.na(arr_delay) & dep_delay < 500) %>% 
  select(tailnum, dep_delay, arr_delay, plane_age) %>% 
  gather(key = type, value = delay, -c(1,4)) %>% 
  ggplot(aes(jitter(plane_age), delay, color = delay)) +
  geom_point(size = 0.2, alpha = 0.1) +
  facet_wrap(type~.) +
  geom_smooth(method = "gam",formula = y ~ s(x, bs = "cs")) +
  xlab("Plane Age") +
  ylab("Delay (in minutes)")
```


# Thursday
Write a function that takes a numeric matrix and transforms each column by substracting by the column mean and dividing by the column standard deviation. Use this function on the state.x77 matrix
```{r}
library(datasets)

my_fn <- function(the_matrix){
  apply(the_matrix, MARGIN = 2, FUN = function(a_col){
    output = (a_col - mean(a_col))/sd(a_col)
    return(output)
  })
}
my_fn(state.x77)
```
Combining apply and lapply write a function to obtain 10 subsamples from state.x77 by selecting 20 rows at random each time, and calculate the column means for each of these subsamples.
```{r}
lapply(1:10, function(X){
  my_sample <- state.x77[sample(nrow(state.x77), 20),]
  apply(my_sample, 2, mean)
})
```
In problem 2, replace lapply by sapply
```{r}
sapply(1:10, function(X){
  my_sample <- state.x77[sample(nrow(state.x77), 20),]
  apply(my_sample, 2, mean)
})
```


