---
title: "week 6"
author: "Mina Mehdinia"
date: '2022-05-09'
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

## In-class Exercises  {.smaller .build}

1. How many flights to Los Angeles (LAX) did each of the legacy carriers (AA, UA, DL or US) have in May from JFK, and what was their average duration?
```{r}
library(nycflights13)
library(dplyr)

flights %>% 
  filter(carrier%in%c("AA", "UA", "DL", "US"),
         month == 5, 
         origin == "JFK",
         dest == "LAX") %>% 
  group_by(carrier) %>%
  summarise(num_flights = n(),
            mean_duration = mean(air_time, na.rm=T))
```


2. What was the shortest flight out of each airport in terms of distance? In terms of duration?

```{r}
#in term of distance
flights %>% 
  group_by(origin) %>% 
  summarise(min_dist = min(distance, na.rm = T))
```

```{r}
#in term of durration
flights %>% 
  group_by(origin) %>% 
  summarise(min_time = min(air_time, na.rm = T))
```


3. Which plane (check the tail number) flew out of each New York airport the most?
```{r}
flights %>% 
  filter(!is.na(tailnum)) %>% 
  group_by(origin, tailnum) %>% 
  tally() %>% 
  filter(n==max(n))

```


<div style="color:forestgreen">
*Optional*

Which date should you fly on if you want to have the lowest possible average departure delay? What about arrival delay? 
</div>
```{r}
#lowest possible average departure delay
flights %>% 
  mutate(date = paste(day, month, year, sep = "/")) %>% 
  group_by(date) %>% 
  summarise(mean.dep.delay = mean(dep_delay, na.rm = T)) %>% 
  ungroup() %>% 
  filter(mean.dep.delay == min(mean.dep.delay)) 

#arrival delay
flights %>% mutate(date = paste(day, month, year, sep= "/")) %>% 
  group_by(date) %>% 
  summarise(mean.dep.delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(mean.dep.delay == min(mean.dep.delay))

```


Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

1. Extract the number of TB cases per country per year
2. Extract the matching population per country per year
3. Divide cases by population, and multiply by 10000
4. Store back in the appropriate place.

```{r}
library(tidyr)
library(dplyr)
table2
cases <- table2 %>% filter(type=="cases") %>% pull(count)
table2 %>% 
  filter(type == "population") %>% 
  mutate(rate = (cases/count)*10000) %>% 
  select(-type, -count)

table4a
table4b

table4a %>% 
  mutate(rate1999=(`1999`/table4b$`1999`)*10000, 
         rate2000=(`2000`/table4b$`2000`)*10000) %>% 
  select(country, rate1999, rate2000)

table1 %>% mutate(rate=(cases/population)*10000)

```

Use tb to generate a tibble having as columns gender and age, and year but restricted to the counts for each year from 1980 through 2008 (i.e., aggregating over countries).

```{r}
library(dplyr)
library(tidyverse)

tb = readr::read_csv(paste0("https://raw.githubusercontent.com/tidyverse/tidyr/main/vignettes/tb.csv")) 

tb %>%
  gather(key = group, value = counts,
         -iso2, -year,
         na.rm = TRUE) %>% 
  separate(col = group, into = c("gender", "age"), sep = 1) %>%
  mutate(
    age = case_when(
      age == "04" ~ "0-4",
      age == "514" ~ "5-14",
      age == "014" ~ "0-14",
      age == "1524" ~ "15-24",
      age == "2534" ~ "25-34",
      age == "3544" ~ "35-44",
      age == "4554" ~ "45-54",
      age == "5564" ~ "55-64",
      age == "65" ~ "65",
      age == "u" ~ NA_character_
      )
  ) %>% 
  spread(key = year, value = counts)


```

