---
title: "Week 8 Problem"
author: "Mina Mehdinia"
date: '2022-05-22'
output: html_document
---
Let’s use the got_chars data set in the repurrrsive package


```{r}
library(datasets)
library(repurrrsive)
library(purrr)
class(got_chars)

```
1)Using the appropriate map function extract the name of all characters in GOT
```{r}
map_chr(got_chars, "name") 
```
2.Use map to extract for all elements of the list the id, name and gender
```{r}
map(got_chars, `[`, c("id", "name", "gender")) 

```
3)Generate a tidy data frame (tibble) containing the variables name, culture, gender, id, born, alive for all characters.
```{r}
map_df(got_chars, `[`, c("name", "culture", "gender", "id", "born", "alive"))
```
Implement arg_max().

Inputs: a function and a vector

Output: the elements of the input where the function returns the highest value.

For example: * arg_max(-10:5, function(x) x ^ 2) should return -10
```{r}
arg_max <- function(f,x){
  x[f(x) == max(f(x))]
}
arg_max(-10:10, f = function(x)x^2)

arg_max(-10:10, f = function(x)sin(x))

sin(-10:10)

```
Use reduce() to find all the elements that appear in at least one entry of the following list
```{r}
(my.list <- map(1:4, ~ sample(1:10, 15, replace = T)))
```
```{r}
reduce(my.list, union)
```
```{r}
accumulate(my.list, union)
```

Rewrite this function as a function factory:
```{r}
boxcox1 <- function(x, lambda) {
  stopifnot(length(lambda) == 1)
  
  if (lambda == 0) {
    log(x)
  } else {
    (x ^ lambda - 1) / lambda
  }
}
```

```{r}
bc.factlambda <- function(lambda){
  stopifnot(length(lambda)== 1)
  force(lambda)
  if(lambda == 0){
    function(x){
      log(x)
    }
  }else{
    function(x){
      (x ^ lambda - 1) / lambda
    }
  }
}

bc.0 <- bc.factlambda(lambda = 0)
bc.other <- bc.factlambda(lambda = 0.3)

x <- rgamma(1000,1,10)

par(mfrow = c(1,3))
hist(x, main = "Originial data")
bc.0(x) %>% hist(main = expression(lambda ==0))
bc.other(x) %>% hist(main = expression(lambda = 0.3))
```
Recall the sleep vs caffeine experiment? Write the function factory my.perm.test, which takes:

1) a data.frame,

2) the name of the variable with the group labels (assume there are only two groups), and

3)the name of the variable with the values, and returns the difference between the means of the permuted groups

Use this function to generate 10,000 draws of the permuted differences for the sleep vs caffeine data found here:
```{r}
Memory <- read.csv("http://www.mosaic-web.org/go/datasets/SleepCaffeine.csv")
```

```{r fig.height=4, fig.width=5}
library(purrr)
library(stringr)
library(ggplot2)
library(tidyverse)

my_perm_wrap <- function(df, label.var, value.var) {
  function() {
    vv <- sample(1:nrow(df), replace = F)
    perm_labels <- df[[label.var]][vv]
    by(df[[value.var]], perm_labels, mean) %>% diff() 
  }
}

my_perm_test <- my_perm_wrap(Memory, label.var = "Group", value.var = "Words")

vec_of_perm_diffs <- map_dbl(1:10000, ~ my_perm_test())

tibble(differ = vec_of_perm_diffs) %>% 
  ggplot(aes(x = differ)) +
  geom_density(fill = "blue") +
  geom_vline(xintercept = (with(Memory, by(Words, Group, mean)) %>%  diff()),
             color = "red", lwd = 2) +
  xlab(expression(mu[caffeine] - mu[sleep])) + 
  labs(title = "Distribution for the permuted differences",
       subtitle = "(between caffeine and sleep treatments)")
```

